---
title: "aspis-chemicals"
description: |
  A short description of the post.
author:
  - name: Thomas Luechtefeld
    url: {}
date: 2022-03-17
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## Introduction
This workflow describes how to retrieve literature for a given compound name and a given query context (query) from the PubMed database. Information on compounds (e.g. its synonyms is retrieved using the PubChem api).

## Packages
```{r}
library(tidyverse)
library(litsearchr)
library(rentrez)
library(easyPubMed)
library(europepmc)
library(readxl)
library(webchem)
# install.packages("googlesheets4")
library(googlesheets4)
```

## Documentation

 - [`{rentrez}`](https://cran.r-project.org/web/packages/rentrez/vignettes/rentrez_tutorial.html#finding-cross-references-entrez_link)
 - [`{webchem}`](https://github.com/ropensci/webchem)
 - 

## List of chemicals 
Here we load the Googlesheet file containing compound info from the ASPIS project cluster for which we will build the query and perform the search. 

see for workflow on how to use R to connect to Googlesheets: https://www.r-bloggers.com/2021/12/how-to-connect-r-to-google-sheets-using-googlesheets4/

To initiate the authentication for your Google account run `gs4_auth()` in the Console. You are presented with the option to select an eaxisting google account or to add a new one. You may choose to link the current session to an already cached token (press enter after selecting the apporpriate Google account), or choose '0' to obtain a new token. 

The Googlesheet we are trying to access here is in 'everybode who has the link has access' mode. So we do not need to authenticate. We can use `gs4_deauth()` here, which in this case is safer to use.

The url to the sheet is stored as a cached value, to not who it here in the code. We use the RStudio API to achieve this


```{r, eval=FALSE}
gs4_deauth()
url_sheet <- rstudioapi::askForSecret(name = "url_sheet")
data_sheet <- read_sheet(url_sheet)
data_sheet
```

## Clean and select data
Let's clean up and reduce the amount of data. Steps to take:

 - Clean up variable names to make them unix-safe
 - The file is not tidy, the columns "Project ID (RISK-HUNT3R)", 
 "Project ID (ONTOX)", "Project ID (PrecisionTox)", "Used in RISK-HUNT3R", 
 "Used in ONTOX", "Used in PrecisionTox" contain variables 
 (`project_id` & `used_in_project`), that are    spread over multiple columns. 
 Here we will not use the information in these columns, so I will remove them

```{r}
data_tidy <- data_sheet |>
  janitor::clean_names() |>
  dplyr::select(-c("project_id_risk_hunt3r":"used_in_precision_tox"))
 data_tidy
```

## Add CID 
CID is the PubChem id. We need this id for other things.
```{r, get_cids_from_names, eval=FALSE}
## get all CID for all chemicals by name
## wrap get_cid() in function with a pause, not to overburden the NCBI server.
get_cid_with_pause <- function(..., wait = 0.5){
  
  x <- get_cid(...)
  profvis::pause(wait)
  return(x)
  
}

hold = 2

data_tidy <- data_tidy |>
  mutate(
    cid = map(
    .x = inchi,
    get_cid_with_pause,
    verbose = TRUE,
    from = 'inchi',
    wait = hold)
  )

data_tidy <- data_tidy |>
  unnest(cid) |>
  dplyr::select(
    compound_name,
    iupac,
    inchi,
    smiles,
    cas_number,
    dtxsid,
    cid
  )
```

## Enrich compound data with synonyms 
See [PUG-REST API PubChem](https://pubchemdocs.ncbi.nlm.nih.gov/pug-rest) for all attributes that can be retrieved. 
```{r, get_synonyms_and_inchi_from_names, eval=FALSE}
get_synonyms_with_pause <- function(..., wait = 0.5){
  
  x <- pc_synonyms(...)
  profvis::pause(wait)
  return(x)
  
}

data_tidy <- data_tidy |>
  mutate(
    synonyms = map(
      cid,
      get_synonyms_with_pause,
      from = "cid",
      wait = hold, 
      verbose = TRUE)) 

data_tidy <- data_tidy |>
  unnest(synonyms) |>
  dplyr::select(
    compound_name,
    synonyms,
    iupac,
    inchi,
    smiles,
    cas_number,
    dtxsid,
    cid
  )

## write as dataset to package
data_ASPIS_with_cid <- data_tidy

## usethis::use_data(data_ASPIS_with_cid)

```

## Result
```{r}

data_ASPIS_with_cid
```


Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


